---
title: "Week5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Spotify Data

Load data from spotify for music through years

```{r carsree}
data <- read.csv('spotify_data_by_year.csv')
data
```

## Lib ggplot load

```{r load libraries}
library('tidyverse')
library('ggplot2')
library('corrplot')
library(dplyr)
str(data)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Plots

```{r plot graphs}
# Danceabilty in years
scatter <- ggplot(data,aes(x=year,y=danceability))+geom_jitter()

scatter +
  labs(x = "Track Year", y = "Dance Hit ") +
  ggtitle("How much dancing people did through time")


# how popularity connects to danceability

```
## Loudness through time
```{r scatter loudness}
scatter2 <- ggplot(data,aes(x=year,y=loudness))+geom_smooth()

scatter2 +
  labs(x = "Track Year", y = "Loudness ") +
  ggtitle("How louder the music through time")

```
## Loudness over years pt2
```{R}
ggplot(data,aes(x = year, y = loudness)) +
   geom_contour(stat = "Identity") +
  ggtitle("Loudness over years")
```

## Correlation Between Variables
```{r correlation}
cor_data <- cor(data)
corrplot(cor_data, method = "color", order = "AOE")
```

## Test Three Variables
```{r multivariate}
ggplot(data, aes(x = year, y = loudness)) +
    geom_tile(aes(color = popularity))
  
```
# Loundness and Popularity
```{R }
ggplot(data,aes(x = year, y = loudness, fill = popularity)) +
   geom_contour(stat = "Identity") +
  ggtitle("Popularity and Loudness over years")
```

# Danceability and Popularity
```{R plot danceability and popularity through time}

ggplot(data,aes(x = year, y = danceability, fill = popularity)) +
  geom_bar(stat = "Identity") +
  ggtitle("Popularity and Danceability over years")
```
## Organize data per decade
```{R }

data_per_dec <- mutate(data,decade = floor(year/10)*10) %>% 
      group_by(decade) %>% 
      summarize_all(mean) %>% 
      select(-year)
data_per_dec
 
```
## Plot data acording decades
```{r}
ggplot(data_per_dec,aes(x=decade,y=acousticness)) +
  geom_contour(stat="Identity")+
  ggtitle("Mean acousticness value per decade")
```

## Mean danceability value 
```{r}
ggplot(data_per_dec,aes(x=decade,y=danceability)) +
  geom_contour(stat="Identity")+
  ggtitle("Mean danceability value per decade")
```
## Median value per attribute
```{r}
data_per_dec2 <- mutate(data,decade = floor(year/10)*10) %>% 
      group_by(decade) %>% 
      summarize_all(median) %>% 
      select(-year)
data_per_dec2
```

## Danceability per decade scale _x_axis
```{r}
ggplot(data_per_dec2,aes(x=decade,y=danceability)) +
  geom_contour(stat="Identity")+
  ggtitle("Median danceability value per decade")+
   scale_x_continuous(breaks= seq(1921, 2021, by=10))
```
```{r}
ggplot(data_per_dec2,aes(x=decade,y=tempo)) +
  geom_contour(stat="Identity")+
  ggtitle("Median tempo value per decade")+
   scale_x_continuous(breaks= seq(1921, 2021, by=10))
```

